\section{Validation}

Validation focuses on whether the programmed model is an adequate
representation of real-world carbon-aware training for the study objectives.
It emphasizes consistency with domain
expectations, rather than low-level code correctness.

\subsection{Structural Trade-off Behaviour}

The first validation checks the core trade-off between CO$_2$ savings and
training time overhead when varying the pause threshold
$\theta_{\text{pause}}$. For fixed region, start time, and model
configuration, stricter thresholds are expected to yield higher relative
emission savings but also higher wall-clock overhead. The simulator is
validated by confirming a monotone trend in this trade-off and by inspecting
the Pareto frontier for reasonable shape and absence of pathological cases,
such as negative savings for very strict thresholds.

\subsection{Cross-Model Scaling}

A second validation examines how emissions and savings scale across different
model architectures, for example DeepSeek V3 and Llama‑3.1 405B. For models
run under the same region and carbon-aware policy, baseline emissions should
approximately scale with total data centre energy, driven by GPU count, power,
GPU-hours and PUE. Similarly, absolute savings should scale with baseline
emissions. Ratios of simulated emissions and savings between models are
compared against ratios implied by their GPU-hours and power. Qualitatively
implausible outcomes, such as a smaller model emitting far more than a larger
one under identical assumptions, would indicate issues.

\subsection{Regional and Temporal Realism}

The third validation targets regional and temporal realism. For each region
and year, an average grid carbon intensity is derived from the input
time-series, and baseline emissions in a no-pause scenario should broadly
follow this ordering for fixed IT energy. Additionally, under a chosen
carbon-aware pausing policy, regions or periods with high temporal variability
in CO$_2$ intensity should exhibit greater potential savings than low-variance,
already clean regions, albeit with more complex pause patterns. Anomalous
results are checked against features of the underlying time-series to ensure
they can be explained by specific training windows.

\subsection{Checkpoint Granularity and Overhead Effects}

Finally, the interaction between pause granularity, checkpoint overhead, and
carbon-aware pausing is validated. Finer pause granularity should allow more
precise avoidance of high-carbon intervals and thus higher potential savings,
but also increase the number of pauses and associated time and energy
overhead. The observed runtime extension and additional energy consumption are
compared to what is implied by the configured per-checkpoint overhead and the
recorded pause count, and the resulting savings–overhead trade-off is checked
for qualitatively reasonable behaviour and diminishing returns at very fine
granularity.

\section{Verification}

Verification ensures that the implementation follows the intended
mathematical model and uses the specified inputs and state updates correctly.
It is carried out via targeted tests of energy and emissions accounting,
pause logic, progress and runtime, data gap handling, and KPI functions.

\subsection{Energy and Emissions Accounting}

The first verification step checks the energy and emissions accounting. At
each time step, IT power, data centre power, and incremental emissions are
computed from GPU counts, training or pause power, PUE and the CO$_2$
intensity. The accumulated emissions over time, as well as baseline and policy
totals, are compared against analytical results in simple synthetic scenarios.
Additionally, the KPI formulas for CO$_2$ savings percentage, time overhead
percentage, and the composite score are verified to match their definitions
for fixed numerical inputs.

\subsection{Pause/Resume Logic and Hysteresis}

The second verification step focuses on the pause/resume logic driven by the
pause threshold $\theta_{\text{pause}}$ and hysteresis
$\delta_{\text{hyst}}$. A synthetic sequence of CO$_2$ intensities is
constructed to trigger known patterns of pausing and resuming according to
the specified hysteresis rule with an implied resume threshold. The simulated
pause state and pause count at each step are compared to the expected
sequence, ensuring that transitions only occur when the intensity crosses the
configured thresholds in the intended direction.

\subsection{Progress and Runtime Consistency}

The third verification step addresses consistency between training progress
and wall-clock time. In non-paused steps, the progress fraction should
increase in proportion to the time step relative to the baseline training
time, while the elapsed wall time grows in every step. In a no-pause
scenario, the simulation is expected to reach full progress exactly at the
baseline runtime within numerical tolerance. In scenarios with pauses, the
simulated policy runtime is compared against the analytical expression
combining baseline time, total paused time, and per-pause checkpoint
overhead, confirming that the runtime extension due to pausing is implemented
correctly.

\subsection{Handling of Data Gaps}

The fourth verification step ensures robust handling of missing or invalid
CO$_2$ data. When the data availability flag indicates missing data, the
simulation must replace the instantaneous intensity with a predefined
region–year average and consistently use this effective value both for
emissions calculations and pause decisions. Synthetic time-series with known
missing points are used to confirm that the effective intensity and resulting
decisions follow this rule at every affected time step.

\subsection{Unit Tests for KPI Functions}

Finally, KPI computations are tested in isolation through unit tests. Fixed
values are supplied for baseline and policy emissions and runtimes, and the
functions computing CO$_2$ savings percentage, time overhead percentage, and
the composite score are checked against manually computed expected results.
This confirms that the KPI implementations are numerically correct,
independent of the rest of the simulation engine.